---
output: html_document
editor_options: 
  chunk_output_type: inline
---
title: "Re-analysis of Differential Expressed Genes"
output:
  html_notebook: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---
Author: Ridzky Yuda, Ph.D.
Datum: 12. Januar 2025
# 1 Installation von die gebrauchte Pakete
```{r}
#remotes::install_github("satijalab/seurat", "seurat5", quiet = TRUE)
#install.packages("sctransform")
#install.packages("harmony")
#BiocManager::install("harmony", version = "3.16")
#BiocManager::install("glmGamPoi")
#BiocManager::install("clustermole")
#install.packages("pathfindR")
#BiocManager::install("dittoSeq")
#BiocManager::install("formatR")
#BiocManager::install("ComplexHeatmap")
#BiocManager::install("fgsea")
#install.packages("msigdbr")
#devtools::install_github("jhrcook/readgmt")
#BiocManager::install("AUCell")
#BiocManager::install("decoupleR")
#BiocManager::install("UCell")
#BiocManager::install("ReactomePA")
#install.packages("ggVennDiagram")

```

# 2 Paket Aktivierung
```{r}
library(ggVennDiagram)
library(ReactomePA)
library(UCell)
library(decoupleR)
library(readgmt)
library(stringr)
library(RColorBrewer)
library(Seurat)
library(patchwork)
library(sctransform)
library(harmony)
library(ggplot2)
library(patchwork)
library(tidyverse)
library(cowplot)
library(dplyr)
library(glmGamPoi)
library(ggsci)
library(RCurl)
library(AnnotationHub)
library(ensembldb)
library(dittoSeq)
library(formatR)
library(ComplexHeatmap)
library(circlize)
library(fgsea)
library(msigdbr)
library(dblyr)
library(SingleR)
library(celldex)
```

# 3 Laden die neueste geklustert Seuratobjek mit neuester Annotation
```{r}
CombinedCluster <- readRDS("~/scratch4-mjung9/ryuda1/RY27_LSK_scRNA-seq_scTE/2024-10-02_Cluster_Annotations/combined3Cluster.Seurat.rds")
```

#4 Fügen die CellType nach Metadaten von Seurat Objekt hinzu
```{r}
labels <-c("LT-HSC", 
           "MPP4/LMPP", 
           "ST-HSC", 
           "CLP", 
           "MPP4/LMPP", 
           "ST-HSC", 
           "CLP", 
           "Unknown", 
           "ST-HSC", 
           "CLP", 
           "MDP", 
           "MPP3", 
           "MPP2",
           "MPP3",
           "B cell prog",
           "Baso-Mast-Prog",
           "ST-HSC",
           "MDP",
           "B cell prog",
           "pDC/DC-prog",
           "T cell prog",
           "ST-HSC",
           "MPP3",
           "Baso-Mast-Prog",
           "T cell prog",
           "MEP",
           "B cell prog",
           "B cell prog",
           "B cell prog",
           "GMP",
           "pDC/DC-prog",
           "CDP")

# create a new column in metadata
CombinedCluster$CellType <- CombinedCluster$seurat_clusters

# replace cluster number by cell type
levels(CombinedCluster$CellType) <-labels
```

#5 Visualize UMAP from the Seurat object
```{r}
#CombinedCluster <- factor(test.data$sample, levels = c("Young control", "Young alcohol", "Old control", "Old alcohol1", "Old alcohol2"))

colors <- c("#e3d132","#99bd4b","#5fc26f","#1F78B4","#A6CEE3", "#E31A1C","#E3751C","#600078", "#33A02C", "#FDBF6F", "#526317", "#ae5e93", "#784620", "#edbbe3", "#eb5252")

All<-print(DimPlot(object = CombinedCluster, 
        reduction = "umap", 
        label = T,
        pt.size = 1,
        repel = T,
        label.size = 4,
        cols = colors)) + ggtitle("Global") +
  theme(plot.title = element_text(hjust=0.5))
All
ggsave("2024-10-04_All_samples_UMAP.pdf", All, device=cairo_pdf,width = 10, height = 8, units = c("in"))
```

#6 Aging Module Score Module Calculation
```{r}
# aging HSC marker genes
aging_markers_table<-readxl::read_xlsx('AgingSignature2.xlsx') %>% as.data.frame() #%>% dplyr::filter(Log2FC > 0)
aging_markers_hsc<-aging_markers_table$Gene

aging_markers_hsc<-aging_markers_hsc[!is.na(aging_markers_hsc)]

CombinedClusterWithAgingModule<- AddModuleScore(CombinedCluster, features=list(aging_markers_hsc), name='Aging_HSC_Marker-Score-Svendsen')

CombinedClusterWithAgingModule@meta.data$group <- factor(x = CombinedClusterWithAgingModule@meta.data$group, levels = c("Young control", "Young alcohol", "Old control", "Old alcohol")) # relevel group order


#YC_with_AS<-subset(CombinedClusterWithAgingModule, group %in% "Young control")
#ycAS<-FeaturePlot(YC_with_AS,
 #           features = "aging_hsc_markers_score1", reduction = "umap",
  #          max.cutoff = 0.3, min.cutoff = -0.4,
   #         pt.size = 0.6,
    #        label = FALSE, repel = TRUE) +
     #       scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

#YA_with_AS<-subset(CombinedClusterWithAgingModule, group %in% "Young alcohol")
#yaAS<-FeaturePlot(YA_with_AS,
 #           features = "aging_hsc_markers_score1", reduction = "umap",
  #          max.cutoff = 0.3, min.cutoff = -0.4,
   #         pt.size = 0.6,
    #        label = FALSE, repel = TRUE) +
     #       scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

#OC_with_AS<-subset(CombinedClusterWithAgingModule, group %in% "Old control")
#ocAS<-FeaturePlot(OC_with_AS,
 #           features = "aging_hsc_markers_score1", reduction = "umap",
  #          max.cutoff = 0.3, min.cutoff = -0.4,
   #         pt.size = 0.6,
    #        label = FALSE, repel = TRUE) +
     #       scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

#OA_with_AS<-subset(CombinedClusterWithAgingModule, group %in% "Old alcohol")
#oaAS<-FeaturePlot(OA_with_AS,
 #           features = "aging_hsc_markers_score1", reduction = "umap",
  #          max.cutoff = 0.3, min.cutoff = -0.4,
   #         pt.size = 0.6,
    #        label = FALSE, repel = TRUE) +
     #       scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")))

LTHSC<-subset(CombinedClusterWithAgingModule, CellType %in% "LT-HSC")
HSC<-subset(CombinedClusterWithAgingModule, CellType %in% c("LT-HSC", "ST-HSC"))

LTHSCviolin<-VlnPlot(LTHSC, features=c('Aging_HSC_Marker-Score-Svendsen1'), group.by = 'group', pt.size=0.3)+stat_summary(fun.y = mean, geom='point', size = 10, colour = "black", shape = 95)

agingscore<-LTHSC@meta.data$`Aging_HSC_Marker-Score-Svendsen1`
group<-LTHSC@meta.data$group

agingscoreextracted<-data.frame(group, agingscore)
xlsx::write.xlsx(agingscoreextracted, "2025-02-7_AgingScore-LT-HSC.xlsx") # save into an Excel file

  

#ggsave("2025-01-30_AgingScoreUMAP_YC.pdf", ycAS, device=cairo_pdf,width = 10, height = 8, units = c("in"))
#ggsave("2025-01-30_AgingScoreUMAP_YA.pdf", yaAS, device=cairo_pdf,width = 10, height = 8, units = c("in"))
#ggsave("2025-01-30_AgingScoreUMAP_OC.pdf", ocAS, device=cairo_pdf,width = 10, height = 8, units = c("in"))
#ggsave("2025-01-30_AgingScoreUMAP_OA.pdf", oaAS, device=cairo_pdf,width = 10, height = 8, units = c("in"))
ggsave("2025-02-07_AgingScoreViolinPlot-LT-HSC_OA.pdf", LTHSCviolin, device=cairo_pdf,width = 10, height = 8, units = c("in"))

```

# UMAP aging scores
```{r}
p1<-FeaturePlot(CombinedClusterWithAgingModule,
            features = "Aging_HSC_Marker-Score-Svendsen1", reduction = "umap",
            max.cutoff = 0.2, min.cutoff = -0.2,
            pt.size = 0.6, 
            split.by = "group",
            label = FALSE, repel = TRUE,
            combine = FALSE) 

fix.sc<-scale_colour_gradientn(colours = rev(brewer.pal(n = 11, name = "RdBu")), limits = c(-0.2, 0.2))

p2 <- lapply(p1, function (x) x + fix.sc)
agingscoreallumap<-CombinePlots(p2)

ggsave("2025-02-07_AgingScoreUMAP-LT-HSC.pdf", agingscoreallumap, device=cairo_pdf,width = 12, height = 10, units = c("in"))

#p1 <- FeaturePlot(pbmc_small, features = c("PPBP", "IGLL5", "SDPR"), combine = FALSE )
#fix.sc <- scale_color_gradientn( colours = c('lightgrey', 'blue'),  limits = c(1, 8))
```



```{r}

#signatures<- list(
  #            HALLMARK_INTERFERON_ALPHA_RESPONSE = #mh_genesets$HALLMARK_INTERFERON_ALPHA_RESPONSE,
   #           HALLMARK_INTERFERON_GAMMA_RESPONSE = #mh_genesets$HALLMARK_INTERFERON_GAMMA_RESPONSE,
    #          HALLMARK_TNFA_SIGNALING_VIA_NFKB = #mh_genesets$HALLMARK_TNFA_SIGNALING_VIA_NFKB,
     #         REACTOME_INTERFERON_SIGNALING = #reactome_genesets$REACTOME_INTERFERON_SIGNALING,
#              REACTOME_TNFR1_INDUCED_NF_KAPPA_B_SIGNALING_PATHWAY = reactome_genesets$REACTOME_TNFR1_INDUCED_NF_KAPPA_B_SIGNALING_PATHWAY,
#              REACTOME_REGULATION_OF_ENDOGENOUS_RETROELEMENTS = reactome_genesets$REACTOME_REGULATION_OF_ENDOGENOUS_RETROELEMENTS,
 #             REACTOME_CELLULAR_SENESCENCE = reactome_genesets$REACTOME_CELLULAR_SENESCENCE,
  #            REACTOME_INNATE_IMMUNE_SYSTEM = reactome_genesets$REACTOME_INNATE_IMMUNE_SYSTEM,
   #           REACTOME_CELL_CYCLE = reactome_genesets$REACTOME_CELL_CYCLE,
    #          REACTOME_EPIGENETIC_REGULATION = reactome_genesets$REACTOME_EPIGENETIC_REGULATION_OF_GENE_EXPRESSION,
     #         GOBP_RESPONSE_TO_BACTERIUM_LIPOPROTEIN = gobp_genesets$GOBP_RESPONSE_TO_BACTERIAL_LIPOPROTEIN,
      #        MARROW_HEMATOPOIETIC_STEM_CELL_AGEING = m8_genesets$TABULA_MURIS_SENIS_MARROW_HEMATOPOIETIC_STEM_CELL_AGEING,
       #        MARROW_HEMATOPOIETIC_PRECURSOR_AGEING = m8_genesets$TABULA_MURIS_SENIS_MARROW_HEMATOPOIETIC_PRECURSOR_CELL_AGEING,
        #      REACTOME_PRC2_METHYLATES_HISTONES_AND_DNA = reactome_genesets$REACTOME_PRC2_METHYLATES_HISTONES_AND_DNA,
         #     RUNX1_MEG_DIFF = reactome_genesets$REACTOME_RUNX1_REGULATES_GENES_INVOLVED_IN_MEGAKARYOCYTE_DIFFERENTIATION_AND_PLATELET_FUNCTION,
          #    DDR_TELOMERE_STRESS = reactome_genesets$REACTOME_DNA_DAMAGE_TELOMERE_STRESS_INDUCED_SENESCENCE
              
              
)

seurat.object<-AddModuleScore_UCell(CombinedClusterWithAgingModule, features = signatures, name = NULL)
```


```{r}
DotPlot<-print(DotPlot(CombinedCluster, cols = "RdYlBu", assay = "RNA", features = c("Procr", "Pdzk1ip1","Ly6a", "Klf1", "Car1", "Dntt", "Cd34", "Satb1", "Mpo", "Calr", "Pf4", "Sdpr", "Cd63", "Cpa3", "Kit"), group.by = "CellType"))+
  theme(axis.text.x=element_text(size=rel(0.8), angle=90))+
  ggtitle("Markers gene expression in different clusters")+
  theme(plot.title = element_text(hjust=0.5))
#ggsave("2024-10-04_Dotplot_LT-HSC_TEs.pdf", DotPlotLTHSC_TE, device=cairo_pdf,width = 15, height = 7, units = c("in"))
```


#8 Differenzielle Expressionsanalyse des LT-HSC-Clusters

#Dataset Vorbereitung vor Gene und TE Expressionsanalyse
```{r}
options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

#Idents(CombinedCluster) <- "seurat_annotations"
CombinedCluster$seurat_annotations <- Idents(CombinedCluster)

# Hier kombiniere ich Old Alcohol RY22 und Old Alcohol 2 RY23 Samples zu Old Alcohol Grup.

CombinedCluster$celltype.stim <- paste(CombinedCluster$seurat_annotations,CombinedCluster$group, sep = "_")
Idents(CombinedCluster) <- "celltype.stim"
CombinedCluster<-PrepSCTFindMarkers(CombinedCluster)

```


#a. LT-HSC DEG Analysis using Wilcoxon
```{r}
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim LT-HSC Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "LT-HSC_Old alcohol", ident.2 = "LT-HSC_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim LT-HSC Cluster
YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "LT-HSC_Young alcohol", ident.2 = "LT-HSC_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim LT-HSC Cluster
OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "LT-HSC_Old control", ident.2 = "LT-HSC_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim LT-HSC Cluster
OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "LT-HSC_Old alcohol", ident.2 = "LT-HSC_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "LT-HSC_Young alcohol", ident.2 = "LT-HSC_Old control", test.use = "wilcox", verbose = FALSE)

OAvsOCWilcox$genes<-rownames(OAvsOCWilcox) # extract gene names and add into a new column
YAvsYCWilcox$genes<-rownames(YAvsYCWilcox) # extract gene names and add into a new column
OCvsYCWilcox$genes<-rownames(OCvsYCWilcox) # extract gene names and add into a new column
OAvsYCWilcox$genes<-rownames(OAvsYCWilcox) # extract gene names and add into a new column
YAvsOCWilcox$genes<-rownames(YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(OAvsOCWilcox, "2025-10-29_OAvsOC_LT-HSC_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(YAvsYCWilcox, "2025-10-29_YAvsYC_LT-HSC_DEG-Wilcox.xlsx")
xlsx::write.xlsx(OCvsYCWilcox, "2025-10-29_OCvsYC_LT-HSC_DEG-Wilcox.xlsx")
xlsx::write.xlsx(OAvsYCWilcox, "2025-10-29_OAvsYC_LT-HSC_DEG-Wilcox.xlsx")
xlsx::write.xlsx(YAvsOCWilcox, "2025-10-29_YAvsOC_LT-HSC_DEG-Wilcox.xlsx")
```

```{r}
# Venn Diagram for differentially expressed genes from LT-HSC OA vs OC and YA vs YC comparisons
# Get significantly differentially expressed gene lists from LT-HSC OA vs OC and ordered based on fold change
Up.OAvsOCWilcox <- OAvsOCWilcox %>%
  dplyr::filter(p_val_adj <= 0.05) %>%
  dplyr::filter(avg_log2FC >= 0.58) %>%
  dplyr::arrange(desc(avg_log2FC))
GenesUpOAvsOC<-Up.OAvsOCWilcox$genes

Dn.OAvsOCWilcox <- OAvsOCWilcox %>%
  dplyr::filter(p_val_adj <= 0.05) %>%
  dplyr::filter(avg_log2FC <= -0.58) %>%
  dplyr::arrange(desc(avg_log2FC))
GenesDnOAvsOC<-Dn.OAvsOCWilcox$genes

Sig.OAvsOCWilcox <- rbind(Up.OAvsOCWilcox, Dn.OAvsOCWilcox)
SigGenesOAvsOC <-Sig.OAvsOCWilcox$genes

# Get significantly differentially expressed gene lists from LT-HSC YA vs YC
Up.YAvsYCWilcox <- YAvsYCWilcox %>%
  dplyr::filter(p_val_adj <= 0.05) %>%
  dplyr::filter(avg_log2FC >= 0.58) %>%
  dplyr::arrange(desc(avg_log2FC))
GenesUpYAvsYC <- Up.YAvsYCWilcox$genes

Dn.YAvsYCWilcox <- YAvsYCWilcox %>%
  dplyr::filter(p_val_adj <= 0.05) %>%
  dplyr::filter(avg_log2FC >= 0.58) %>%
  dplyr::arrange(desc(avg_log2FC))
GenesDnYAvsYC <- Dn.YAvsYCWilcox$genes

Sig.YAvsYCWilcox <-rbind(Up.YAvsYCWilcox, Dn.YAvsYCWilcox)
SigGenesYAvsYC <-Sig.YAvsYCWilcox$genes

# Get significantly differentially expressed gene lists from LT-HSC OC vs YC
Up.OCvsYCWilcox <- OCvsYCWilcox %>%
  dplyr::filter(p_val_adj <= 0.05) %>%
  dplyr::filter(avg_log2FC >= 0.58) %>%
  dplyr::arrange(desc(avg_log2FC))
GenesUpOCvsYC <- Up.OCvsYCWilcox$genes

Dn.OCvsYCWilcox <- OCvsYCWilcox %>%
  dplyr::filter(p_val_adj <= 0.05) %>%
  dplyr::filter(avg_log2FC >= 0.58) %>%
  dplyr::arrange(desc(avg_log2FC))
GenesDnOCvsYC <- Dn.OCvsYCWilcox$genes

Sig.OCvsYCWilcox <-rbind(Up.OCvsYCWilcox, Dn.OCvsYCWilcox)
SigGenesOCvsYC <-Sig.OCvsYCWilcox$genes

# VennDiagram Upregulated genes
gene_list_up <- list(A = sample(GenesUpYAvsYC),
                  B = sample(GenesUpOAvsOC))

up<-ggVennDiagram(gene_list_up,
              category.names = c("Up in YA vs YC", "Up in OA vs OC"), label = "count") + scale_x_continuous(expand = expansion(mult = .2)) + scale_fill_distiller(palette = "Reds", direction = 1) + coord_flip()
ggsave("2025-06-09_VennDiagram_LT-HSC_Upregulated_genes.pdf", up, device=cairo_pdf,width = 5, height = 4, units = c("in"))

# find overlaps and unique genes in the VennDiagram of upregulated genes
intersect(GenesUpYAvsYC, GenesUpOAvsOC) # genes upregulated due to EtOH in both Y and O
setdiff(GenesUpYAvsYC, GenesUpOAvsOC) # genes upregulated due to EtOH only in Young
setdiff(GenesUpOAvsOC, GenesUpYAvsYC) # genes upregulated due to EtOH only in Old

# Plot VennDiagram of Downregulated genes
gene_list_dn <- list(D = sample(GenesDnYAvsYC),
                  E = sample(GenesDnOAvsOC))

dn<-ggVennDiagram(gene_list_dn,
              category.names = c("Down in YA vs YC", "Down in OA vs OC"), label = "count") + scale_x_continuous(expand = expansion(mult = .2)) + scale_fill_distiller(palette = "Reds", direction = 1) + coord_flip()
ggsave("2025-06-09_VennDiagram_LT-HSC_Downregulated_genes.pdf", dn, device=cairo_pdf,width = 5, height = 4, units = c("in"))
# find overlaps and unique genes in the VennDiagram of downregulated genes
intersect(GenesDnYAvsYC, GenesDnOAvsOC) # genes downregulated due to EtOH in both Y and O
setdiff(GenesDnYAvsYC, GenesDnOAvsOC) # genes downregulated due to EtOH only in Young
setdiff(GenesDnOAvsOC, GenesDnYAvsYC) # genes downregulated due to EtOH only in Old


```

```{r}
genes <- paste0("gene",1:1000)
set.seed(20210302)
gene_list <- list(A = sample(genes,100),
                  B = sample(genes,200),
                  C = sample(genes,300),
                  D = sample(genes,200))

library(ggVennDiagram)
library(ggplot2)
```

# b. MPP3 DEG Analysis using Wilcoxon 
```{r}
#MPP3 DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim MPP3 Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

MPP3_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP3_Old alcohol", ident.2 = "MPP3_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim MPP3Cluster
MPP3_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "MPP3_Young alcohol", ident.2 = "MPP3_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim MPP3 Cluster
MPP3_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP3_Old control", ident.2 = "MPP3_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim MPP3 Cluster
MPP3_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP3_Old alcohol", ident.2 = "MPP3_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
MPP3_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP3_Young alcohol", ident.2 = "MPP3_Old control", test.use = "wilcox", verbose = FALSE)

MPP3_OAvsOCWilcox$genes<-rownames(MPP3_OAvsOCWilcox) # extract gene names and add into a new column
MPP3_YAvsYCWilcox$genes<-rownames(MPP3_YAvsYCWilcox) # extract gene names and add into a new column
MPP3_OCvsYCWilcox$genes<-rownames(MPP3_OCvsYCWilcox) # extract gene names and add into a new column
MPP3_OAvsYCWilcox$genes<-rownames(MPP3_OAvsYCWilcox) # extract gene names and add into a new column
MPP3_YAvsOCWilcox$genes<-rownames(MPP3_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(MPP3_OAvsOCWilcox, "2025-01-15_OAvsOC_MPP3_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(MPP3_YAvsYCWilcox, "2025-01-15_YAvsYC_MPP3_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP3_OCvsYCWilcox, "2025-01-15_OCvsYC_MPP3_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP3_OAvsYCWilcox, "2025-01-15_OAvsYC_MPP3_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP3_YAvsOCWilcox, "2025-01-15_YAvsOC_MPP3_DEG-Wilcox.xlsx")
```

# c. GMP DEG Analysis using Wilcoxon
```{r}
#GMP DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim GMP Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

GMP_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "GMP_Old alcohol", ident.2 = "GMP_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim GMP Cluster
GMP_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "GMP_Young alcohol", ident.2 = "GMP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim GMP Cluster
GMP_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "GMP_Old control", ident.2 = "GMP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim GMP Cluster
GMP_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "GMP_Old alcohol", ident.2 = "GMP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
GMP_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "GMP_Young alcohol", ident.2 = "GMP_Old control", test.use = "wilcox", verbose = FALSE)

GMP_OAvsOCWilcox$genes<-rownames(GMP_OAvsOCWilcox) # extract gene names and add into a new column
GMP_YAvsYCWilcox$genes<-rownames(GMP_YAvsYCWilcox) # extract gene names and add into a new column
GMP_OCvsYCWilcox$genes<-rownames(GMP_OCvsYCWilcox) # extract gene names and add into a new column
GMP_OAvsYCWilcox$genes<-rownames(GMP_OAvsYCWilcox) # extract gene names and add into a new column
GMP_YAvsOCWilcox$genes<-rownames(GMP_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(GMP_OAvsOCWilcox, "2025-01-15_OAvsOC_GMP_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(GMP_YAvsYCWilcox, "2025-01-15_YAvsYC_GMP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(GMP_OCvsYCWilcox, "2025-01-15_OCvsYC_GMP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(GMP_OAvsYCWilcox, "2025-01-15_OAvsYC_GMP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(GMP_YAvsOCWilcox, "2025-01-15_YAvsOC_GMP_DEG-Wilcox.xlsx")
```

# d. MEP DEG Analysis using Wilcoxon
```{r}
#MEP DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim MEP Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

MEP_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MEP_Old alcohol", ident.2 = "MEP_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim MEP Cluster
MEP_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "MEP_Young alcohol", ident.2 = "MEP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim MEP Cluster
MEP_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MEP_Old control", ident.2 = "MEP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim MEP Cluster
MEP_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MEP_Old alcohol", ident.2 = "MEP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
MEP_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MEP_Young alcohol", ident.2 = "MEP_Old control", test.use = "wilcox", verbose = FALSE)

MEP_OAvsOCWilcox$genes<-rownames(MEP_OAvsOCWilcox) # extract gene names and add into a new column
MEP_YAvsYCWilcox$genes<-rownames(MEP_YAvsYCWilcox) # extract gene names and add into a new column
MEP_OCvsYCWilcox$genes<-rownames(MEP_OCvsYCWilcox) # extract gene names and add into a new column
MEP_OAvsYCWilcox$genes<-rownames(MEP_OAvsYCWilcox) # extract gene names and add into a new column
MEP_YAvsOCWilcox$genes<-rownames(MEP_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(MEP_OAvsOCWilcox, "2025-01-15_OAvsOC_MEP_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(MEP_YAvsYCWilcox, "2025-01-15_YAvsYC_MEP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MEP_OCvsYCWilcox, "2025-01-15_OCvsYC_MEP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MEP_OAvsYCWilcox, "2025-01-15_OAvsYC_MEP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MEP_YAvsOCWilcox, "2025-01-15_YAvsOC_MEP_DEG-Wilcox.xlsx")
```


# e. MPP2 DEG Analysis using Wilcoxon
```{r}
#MPP2 DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim MPP2 Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

MPP2_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP2_Old alcohol", ident.2 = "MPP2_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim MPP2 Cluster
MPP2_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "MPP2_Young alcohol", ident.2 = "MPP2_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim MPP2 Cluster
MPP2_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP2_Old control", ident.2 = "MPP2_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim MPP2 Cluster
MPP2_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP2_Old alcohol", ident.2 = "MPP2_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
MPP2_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP2_Young alcohol", ident.2 = "MPP2_Old control", test.use = "wilcox", verbose = FALSE)

MPP2_OAvsOCWilcox$genes<-rownames(MPP2_OAvsOCWilcox) # extract gene names and add into a new column
MPP2_YAvsYCWilcox$genes<-rownames(MPP2_YAvsYCWilcox) # extract gene names and add into a new column
MPP2_OCvsYCWilcox$genes<-rownames(MPP2_OCvsYCWilcox) # extract gene names and add into a new column
MPP2_OAvsYCWilcox$genes<-rownames(MPP2_OAvsYCWilcox) # extract gene names and add into a new column
MPP2_YAvsOCWilcox$genes<-rownames(MPP2_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(MPP2_OAvsOCWilcox, "2025-01-15_OAvsOC_MPP2_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(MPP2_YAvsYCWilcox, "2025-01-15_YAvsYC_MPP2_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP2_OCvsYCWilcox, "2025-01-15_OCvsYC_MPP2_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP2_OAvsYCWilcox, "2025-01-15_OAvsYC_MPP2_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP2_YAvsOCWilcox, "2025-01-15_YAvsOC_MPP2_DEG-Wilcox.xlsx")
```
# f. MPP4 DEG Analysis using Wilcoxon
```{r}
#MPP4 DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim MPP4 Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

MPP4_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP4_Old alcohol", ident.2 = "MPP4_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim MPP4 Cluster
MPP4_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "MPP4_Young alcohol", ident.2 = "MPP4_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim MPP4 Cluster
MPP4_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP4_Old control", ident.2 = "MPP4_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim MPP4 Cluster
MPP4_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP4_Old alcohol", ident.2 = "MPP4_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
MPP4_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MPP4_Young alcohol", ident.2 = "MPP4_Old control", test.use = "wilcox", verbose = FALSE)

MPP4_OAvsOCWilcox$genes<-rownames(MPP4_OAvsOCWilcox) # extract gene names and add into a new column
MPP4_YAvsYCWilcox$genes<-rownames(MPP4_YAvsYCWilcox) # extract gene names and add into a new column
MPP4_OCvsYCWilcox$genes<-rownames(MPP4_OCvsYCWilcox) # extract gene names and add into a new column
MPP4_OAvsYCWilcox$genes<-rownames(MPP4_OAvsYCWilcox) # extract gene names and add into a new column
MPP4_YAvsOCWilcox$genes<-rownames(MPP4_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(MPP4_OAvsOCWilcox, "2025-01-15_OAvsOC_MPP4_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(MPP4_YAvsYCWilcox, "2025-01-15_YAvsYC_MPP4_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP4_OCvsYCWilcox, "2025-01-15_OCvsYC_MPP4_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP4_OAvsYCWilcox, "2025-01-15_OAvsYC_MPP4_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MPP4_YAvsOCWilcox, "2025-01-15_YAvsOC_MPP4_DEG-Wilcox.xlsx")
```

# g. ST-HSC DEG Analysis using Wilcoxon
```{r}
#ST-HSC DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim ST-HSC Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

ST_HSC_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "ST-HSC_Old alcohol", ident.2 = "ST-HSC_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim ST-HSC Cluster
ST_HSC_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "ST-HSC_Young alcohol", ident.2 = "ST-HSC_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim ST-HSC Cluster
ST_HSC_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "ST-HSC_Old control", ident.2 = "ST-HSC_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim ST-HSC Cluster
ST_HSC_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "ST-HSC_Old alcohol", ident.2 = "ST-HSC_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
ST_HSC_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "ST-HSC_Young alcohol", ident.2 = "ST-HSC_Old control", test.use = "wilcox", verbose = FALSE)

ST_HSC_OAvsOCWilcox$genes<-rownames(ST_HSC_OAvsOCWilcox) # extract gene names and add into a new column
ST_HSC_YAvsYCWilcox$genes<-rownames(ST_HSC_YAvsYCWilcox) # extract gene names and add into a new column
ST_HSC_OCvsYCWilcox$genes<-rownames(ST_HSC_OCvsYCWilcox) # extract gene names and add into a new column
ST_HSC_OAvsYCWilcox$genes<-rownames(ST_HSC_OAvsYCWilcox) # extract gene names and add into a new column
ST_HSC_YAvsOCWilcox$genes<-rownames(ST_HSC_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(ST_HSC_OAvsOCWilcox, "2025-01-15_OAvsOC_ST-HSC_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(ST_HSC_YAvsYCWilcox, "2025-01-15_YAvsYC_ST-HSC_DEG-Wilcox.xlsx")
xlsx::write.xlsx(ST_HSC_OCvsYCWilcox, "2025-01-15_OCvsYC_ST-HSC_DEG-Wilcox.xlsx")
xlsx::write.xlsx(ST_HSC_OAvsYCWilcox, "2025-01-15_OAvsYC_ST-HSC_DEG-Wilcox.xlsx")
xlsx::write.xlsx(ST_HSC_YAvsOCWilcox, "2025-01-15_YAvsOC_ST-HSC_DEG-Wilcox.xlsx")
```

# h. MDP DEG Analysis using Wilcoxon
```{r}
#MDP DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim MDP Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

MDP_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MDP_Old alcohol", ident.2 = "MDP_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim MDP Cluster
MDP_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "MDP_Young alcohol", ident.2 = "MDP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim MDP Cluster
MDP_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MDP_Old control", ident.2 = "MDP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim MDP Cluster
MDP_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MDP_Old alcohol", ident.2 = "MDP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
MDP_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "MDP_Young alcohol", ident.2 = "MDP_Old control", test.use = "wilcox", verbose = FALSE)

MDP_OAvsOCWilcox$genes<-rownames(MDP_OAvsOCWilcox) # extract gene names and add into a new column
MDP_YAvsYCWilcox$genes<-rownames(MDP_YAvsYCWilcox) # extract gene names and add into a new column
MDP_OCvsYCWilcox$genes<-rownames(MDP_OCvsYCWilcox) # extract gene names and add into a new column
MDP_OAvsYCWilcox$genes<-rownames(MDP_OAvsYCWilcox) # extract gene names and add into a new column
MDP_YAvsOCWilcox$genes<-rownames(MDP_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(MDP_OAvsOCWilcox, "2025-01-15_OAvsOC_MDP_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(MDP_YAvsYCWilcox, "2025-01-15_YAvsYC_MDP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MDP_OCvsYCWilcox, "2025-01-15_OCvsYC_MDP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MDP_OAvsYCWilcox, "2025-01-15_OAvsYC_MDP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(MDP_YAvsOCWilcox, "2025-01-15_YAvsOC_MDP_DEG-Wilcox.xlsx")
```

# i. CLP DEG Analysis using Wilcoxon
```{r}
#CLP DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim CLP Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

CLP_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CLP_Old alcohol", ident.2 = "CLP_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim CLP Cluster
CLP_HSC_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "CLP_Young alcohol", ident.2 = "CLP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim CLP Cluster
CLP_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CLP_Old control", ident.2 = "CLP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim CLP Cluster
CLP_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CLP_Old alcohol", ident.2 = "CLP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
CLP_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CLP_Young alcohol", ident.2 = "CLP_Old control", test.use = "wilcox", verbose = FALSE)

CLP_OAvsOCWilcox$genes<-rownames(CLP_OAvsOCWilcox) # extract gene names and add into a new column
CLP_HSC_YAvsYCWilcox$genes<-rownames(CLP_HSC_YAvsYCWilcox) # extract gene names and add into a new column
CLP_OCvsYCWilcox$genes<-rownames(CLP_OCvsYCWilcox) # extract gene names and add into a new column
CLP_OAvsYCWilcox$genes<-rownames(CLP_OAvsYCWilcox) # extract gene names and add into a new column
CLP_YAvsOCWilcox$genes<-rownames(CLP_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(CLP_OAvsOCWilcox, "2025-01-15_OAvsOC_CLP_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(CLP_HSC_YAvsYCWilcox, "2025-01-15_YAvsYC_CLP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(CLP_OCvsYCWilcox, "2025-01-15_OCvsYC_CLP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(CLP_OAvsYCWilcox, "2025-01-15_OAvsYC_CLP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(CLP_YAvsOCWilcox, "2025-01-15_YAvsOC_CLP_DEG-Wilcox.xlsx")
```


# j. CDP DEG Analysis using Wilcoxon
```{r}
#CDP DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim CDP Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

CDP_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CDP_Old alcohol", ident.2 = "CDP_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim CDP Cluster
CDP_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "CDP_Young alcohol", ident.2 = "CDP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim CDP Cluster
CDP_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CDP_Old control", ident.2 = "CDP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim CDP Cluster
CDP_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CDP_Old alcohol", ident.2 = "CDP_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
CDP_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "CDP_Young alcohol", ident.2 = "CDP_Old control", test.use = "wilcox", verbose = FALSE)

CDP_OAvsOCWilcox$genes<-rownames(CDP_OAvsOCWilcox) # extract gene names and add into a new column
CDP_YAvsYCWilcox$genes<-rownames(CDP_YAvsYCWilcox) # extract gene names and add into a new column
CDP_OCvsYCWilcox$genes<-rownames(CDP_OCvsYCWilcox) # extract gene names and add into a new column
CDP_OAvsYCWilcox$genes<-rownames(CDP_OAvsYCWilcox) # extract gene names and add into a new column
CDP_YAvsOCWilcox$genes<-rownames(CDP_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(CDP_OAvsOCWilcox, "2025-01-15_OAvsOC_CDP_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(CDP_YAvsYCWilcox, "2025-01-15_YAvsYC_CDP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(CDP_OCvsYCWilcox, "2025-01-15_OCvsYC_CDP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(CDP_OAvsYCWilcox, "2025-01-15_OAvsYC_CDP_DEG-Wilcox.xlsx")
xlsx::write.xlsx(CDP_YAvsOCWilcox, "2025-01-15_YAvsOC_CDP_DEG-Wilcox.xlsx")
```


# k. B cell Prog DEG Analysis using Wilcoxon
```{r}
#B cell pro DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim B cell pro Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

Bcellpro_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "B cell prog_Old alcohol", ident.2 = "B cell prog_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim Bcellpro Cluster
Bcellpro_HSC_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "B cell prog_Young alcohol", ident.2 = "B cell prog_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim Bcellpro Cluster
Bcellpro_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "B cell prog_Old control", ident.2 = "B cell prog_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim Bcellpro Cluster
Bcellpro_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "B cell prog_Old alcohol", ident.2 = "B cell prog_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
Bcellpro_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "B cell prog_Young alcohol", ident.2 = "B cell prog_Old control", test.use = "wilcox", verbose = FALSE)

Bcellpro_OAvsOCWilcox$genes<-rownames(Bcellpro_OAvsOCWilcox) # extract gene names and add into a new column
Bcellpro_HSC_YAvsYCWilcox$genes<-rownames(Bcellpro_HSC_YAvsYCWilcox) # extract gene names and add into a new column
Bcellpro_OCvsYCWilcox$genes<-rownames(Bcellpro_OCvsYCWilcox) # extract gene names and add into a new column
Bcellpro_OAvsYCWilcox$genes<-rownames(Bcellpro_OAvsYCWilcox) # extract gene names and add into a new column
Bcellpro_YAvsOCWilcox$genes<-rownames(Bcellpro_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(Bcellpro_OAvsOCWilcox, "2025-01-15_OAvsOC_Bcellpro_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(Bcellpro_HSC_YAvsYCWilcox, "2025-01-15_YAvsYC_Bcellpro_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Bcellpro_OCvsYCWilcox, "2025-01-15_OCvsYC_Bcellpro_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Bcellpro_OAvsYCWilcox, "2025-01-15_OAvsYC_Bcellpro_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Bcellpro_YAvsOCWilcox, "2025-01-15_YAvsOC_Bcellpro_DEG-Wilcox.xlsx")
```


# l. T cell Progenitor DEG Analysis using Wilcoxon
```{r}
#T cell pro DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim T cell pro Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

Tcellpro_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "T cell progenitor_Old alcohol", ident.2 = "T cell progenitor_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim Tcellpro Cluster
Tcellpro_HSC_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "T cell progenitor_Young alcohol", ident.2 = "T cell progenitor_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim Tcellpro Cluster
Tcellpro_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "T cell progenitor_Old control", ident.2 = "T cell progenitor_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim Tcellpro Cluster
Tcellpro_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "T cell progenitor_Old alcohol", ident.2 = "T cell progenitor_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
Tcellpro_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "T cell progenitor_Young alcohol", ident.2 = "T cell progenitor_Old control", test.use = "wilcox", verbose = FALSE)

Tcellpro_OAvsOCWilcox$genes<-rownames(Tcellpro_OAvsOCWilcox) # extract gene names and add into a new column
Tcellpro_YAvsYCWilcox$genes<-rownames(Tcellpro_YAvsYCWilcox) # extract gene names and add into a new column
Tcellpro_OCvsYCWilcox$genes<-rownames(Tcellpro_OCvsYCWilcox) # extract gene names and add into a new column
Tcellpro_OAvsYCWilcox$genes<-rownames(Tcellpro_OAvsYCWilcox) # extract gene names and add into a new column
Tcellpro_YAvsOCWilcox$genes<-rownames(Tcellpro_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(Tcellpro_OAvsOCWilcox, "2025-01-15_OAvsOC_Tcellpro_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(Tcellpro_YAvsYCWilcox, "2025-01-15_YAvsYC_Tcellpro_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Tcellpro_OCvsYCWilcox, "2025-01-15_OCvsYC_Tcellpro_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Tcellpro_OAvsYCWilcox, "2025-01-15_OAvsYC_Tcellpro_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Tcellpro_YAvsOCWilcox, "2025-01-15_YAvsOC_Tcellpro_DEG-Wilcox.xlsx")
```


# m. BASO/Mast progenitor Progenitor DEG Analysis using Wilcoxon
```{r}
#BASO/Mast progenitor DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim BASO/Mast progenitor Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

BasoMastProg_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Baso/Mast progenitor_Old alcohol", ident.2 = "Baso/Mast progenitor_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim BASO/Mast progenitor Cluster
BasoMastProg_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "Baso/Mast progenitor_Young alcohol", ident.2 = "Baso/Mast progenitor_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim BASO/Mast progenitor Cluster
BasoMastProg_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Baso/Mast progenitor_Old control", ident.2 = "Baso/Mast progenitor_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim BASO/Mast progenitor Cluster
BasoMastProg_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Baso/Mast progenitor_Old alcohol", ident.2 = "Baso/Mast progenitor_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
BasoMastProg_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Baso/Mast progenitor_Young alcohol", ident.2 = "Baso/Mast progenitor_Old control", test.use = "wilcox", verbose = FALSE)

BasoMastProg_OAvsOCWilcox$genes<-rownames(BasoMastProg_OAvsOCWilcox) # extract gene names and add into a new column
BasoMastProg_YAvsYCWilcox$genes<-rownames(BasoMastProg_YAvsYCWilcox) # extract gene names and add into a new column
BasoMastProg_OCvsYCWilcox$genes<-rownames(BasoMastProg_OCvsYCWilcox) # extract gene names and add into a new column
BasoMastProg_OAvsYCWilcox$genes<-rownames(BasoMastProg_OAvsYCWilcox) # extract gene names and add into a new column
BasoMastProg_YAvsOCWilcox$genes<-rownames(BasoMastProg_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(BasoMastProg_OAvsOCWilcox, "2025-01-15_OAvsOC_BasoMastProg_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(BasoMastProg_YAvsYCWilcox, "2025-01-15_YAvsYC_BasoMastProg_DEG-Wilcox.xlsx")
xlsx::write.xlsx(BasoMastProg_OCvsYCWilcox, "2025-01-15_OCvsYC_BasoMastProgDEG-Wilcox.xlsx")
xlsx::write.xlsx(BasoMastProg_OAvsYCWilcox, "2025-01-15_OAvsYC_BasoMastProg_DEG-Wilcox.xlsx")
xlsx::write.xlsx(BasoMastProg_YAvsOCWilcox, "2025-01-15_YAvsOC_BasoMastProg_DEG-Wilcox.xlsx")
```


# n. Unknown progenitor Progenitor DEG Analysis using Wilcoxon
```{r}
#Unknown progenitor DEG Analysis using Wilcoxon
# Hier analyziere ich die DEGs von Old Alcohol gegen Old Control beim Unknown progenitor Cluster mit Wilcoxon-Rank Sum Test

options(future.globals.maxSize = 26000 * 1024^2) # Diese Linie ist fur den Arbeitspeicher RAM vergrossern. Andernsfall das Kode arbeitet nicht

Unknown_OAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Unknown_Old alcohol", ident.2 = "Unknown_Old control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Young Alcohol gegen Young control beim Unknown progenitor Cluster
Unknown_YAvsYCWilcox <- FindMarkers(CombinedCluster, ident.1 = "Unknown_Young alcohol", ident.2 = "Unknown_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Control gegen Young Control beim Unknown progenitor Cluster
Unknown_OCvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Unknown_Old control", ident.2 = "Unknown_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von Old Alcohol gegen Young Control beim Unknown progenitor Cluster
Unknown_OAvsYCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Unknown_Old alcohol", ident.2 = "Unknown_Young control", test.use = "wilcox", verbose = FALSE)

# Hier analyziere ich DEGs von YA gegen OC group
Unknown_YAvsOCWilcox<- FindMarkers(CombinedCluster, ident.1 = "Unknown_Young alcohol", ident.2 = "Unknown_Old control", test.use = "wilcox", verbose = FALSE)

Unknown_OAvsOCWilcox$genes<-rownames(Unknown_OAvsOCWilcox) # extract gene names and add into a new column
Unknown_YAvsYCWilcox$genes<-rownames(Unknown_YAvsYCWilcox) # extract gene names and add into a new column
Unknown_OCvsYCWilcox$genes<-rownames(Unknown_OCvsYCWilcox) # extract gene names and add into a new column
Unknown_OAvsYCWilcox$genes<-rownames(Unknown_OAvsYCWilcox) # extract gene names and add into a new column
Unknown_YAvsOCWilcox$genes<-rownames(Unknown_YAvsOCWilcox) # extract gene names and add into a new column
xlsx::write.xlsx(Unknown_OAvsOCWilcox, "2025-01-15_OAvsOC_Unknown_DEG-Wilcox.xlsx") # save into an Excel file
xlsx::write.xlsx(Unknown_YAvsYCWilcox, "2025-01-15_YAvsYC_Unknown_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Unknown_OCvsYCWilcox, "2025-01-15_OCvsYC_Unknown_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Unknown_OAvsYCWilcox, "2025-01-15_OAvsYC_Unknown_DEG-Wilcox.xlsx")
xlsx::write.xlsx(Unknown_YAvsOCWilcox, "2025-01-15_YAvsOC_Unknown_DEG-Wilcox.xlsx")
```

#9 Volcano plot visualization
# LT-HSC volcano plot
```{r}
# Add a column to indicate significantly differentially expressed genes
#data <- ST_HSC_OAvsOCWilcox %>% 
 # mutate(significant = ifelse(abs(avg_log2FC) > 0.6 & p_val_adj < 0.01, "Significant", "Not significant"))

#top 50 DEGs OA vs OC ST-HSC
#top_genes <- data %>% 
 # dplyr::filter(p_val_adj < 0.01, abs(avg_log2FC) > 0.6) %>% 
  #arrange(p_val_adj) %>% 
  #slice_head(n = 100) %>% 
  #pull(genes)

#=============== plot DEG OA vs OC#====================
# Plot using EnhancedVolcano
LTHSC_OAvsOC<-EnhancedVolcano(OAvsOCWilcox,
    lab = OAvsOCWilcox$genes,
    selectLab = c("Cdkn1a", "Fos", "Jun", 
                              "Egr1", "Klf2", "Klf4", "Mpo", "Ctsg", "Irf8", "Cd74", "Ighv1-85", "Igkv3-4","Ly6d", "Hoxb8", "Mir5107",  "Rtl8c", "Matn4", "H2-Q4", "Jun", "Junb", "Cand2", "Rn18s-rs5", "Snora62", "mt-Cytb", "Nfkbiz", "Isg15", "Ifitm1", "Ifi27", "Ifi27l2a", "Ifi44"),
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-8, 8),
    ylim = c(0, 320),
    title = 'DEGs of Alcohol vs Control in old mouse LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.5,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 3.0,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

ggsave("2025-02-20_Volcano_LT-HSC_OAvsOC.pdf", LTHSC_OAvsOC, device=cairo_pdf,width = 10, height = 7, units = c("in"))

#=============== plot DEG YA vs YC#====================
# Plot using EnhancedVolcano
LTHSC_YAvsYC<-EnhancedVolcano(YAvsYCWilcox,
    lab = YAvsYCWilcox$genes,
    selectLab = c("Cdkn1a", "Fos",
                              "Egr1", "Klf2", "Cd74", "Hoxb8", "H2-Q4", "Junb", "Cand2", "Snora62", "mt-Cytb", "Nfkbid", "Ifitm1"),
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-8, 8),
    ylim = c(0, 320),
    title = 'DEGs of Alcohol vs Control in young mouse LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.5,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 3.0,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

ggsave("2025-02-20_Volcano_LT-HSC_YAvsYC.pdf", LTHSC_YAvsYC, device=cairo_pdf,width = 10, height = 7, units = c("in"))

#=============== plot DEG OC vs YC#====================
# Make sure gene labels are in a column named 'genes'
OCvsYCWilcox$genes <- rownames(OCvsYCWilcox)

# Create volcano plot
LTHSC_OCvsYC <- EnhancedVolcano(OCvsYCWilcox,
  lab = OCvsYCWilcox$genes,
  selectLab = c("Cdkn1a", "Fos", "Egr1", "Klf2", "Cd74", 
                "Hoxb8", "H2-Q4", "Junb", "Cand2", "Snora62", 
                "mt-Cytb", "Nfkbid", "Ifitm1"),
  x = 'avg_log2FC',
  y = 'p_val_adj',
  xlim = c(-8, 8),
  ylim = c(0, 320),
  title = 'DEGs of Old vs Young Control in LT-HSC',
  xlab = bquote(~Log[2]~ 'fold change'),
  pCutoff = 0.05,
  FCcutoff = 0.5,
  cutoffLineType = 'twodash',
  cutoffLineWidth = 0.8,
  pointSize = 1.0,
  labSize = 3.0,
  colAlpha = 1,
  legendLabels = c('Not sig.', 'Log2FC', 'p-value', 'p-value & Log2FC'),
  legendPosition = 'right',
  legendLabSize = 12,
  legendIconSize = 5.0,
  drawConnectors = TRUE,
  max.overlaps = 100,
  gridlines.minor = FALSE
)

# Save the plot
ggsave("2025-02-20_Volcano_LT-HSC_OCvsYC.pdf", LTHSC_OCvsYC, device = cairo_pdf, width = 10, height = 7, units = "in")

```

# ST-HSC volcano plot
```{r}
# Add a column to indicate significantly differentially expressed genes
#data <- ST_HSC_OAvsOCWilcox %>% 
 # mutate(significant = ifelse(abs(avg_log2FC) > 0.6 & p_val_adj < 0.01, "Significant", "Not significant"))

#top 50 DEGs OA vs OC ST-HSC
#top_genes <- data %>% 
 # dplyr::filter(p_val_adj < 0.01, abs(avg_log2FC) > 0.6) %>% 
  #arrange(p_val_adj) %>% 
  #slice_head(n = 100) %>% 
  #pull(genes)

# Plot using EnhancedVolcano
EnhancedVolcano(ST_HSC_OAvsOCWilcox,
    lab = ST_HSC_OAvsOCWilcox$genes,
    selectLab = c("Cdkn1a", "Fos", "Jun", 
                              "Egr1", "Klf2", "Klf4", "Mpo", "Ctsg", "Irf8", "Mki67", "Siglece", "Siglecf", "Ly86", "Fcgr3", "Cd74","H2-T3", "Ighv1-85", "Igkv3-4","Cebpa", "Ly6d","Cdk1", "Csf1r", "Csf2rb", "Hoxb8", "Hoxb9", "Mir5107", "Il1rapl2", "Rtl8c", "Matn4", "Mpl", "H2-Q4", "Irf6", "Cavin2", "Neo1", "Wwtr1", "Hoxb2", "Hoxb6", "Cdkn1c"),
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-8, 8),
    ylim = c(0, 320),
    title = 'DEGs of Alcohol vs Control in old mouse ST-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 10e-2,
    FCcutoff = 0.5,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 4.0,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 40,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


```


# 10. Gene set enrichment analysis
# a. load downloaded genesets from MsigDB
```{r}
aging_signature_HSC<- list(aging_HSC = aging_markers_hsc)

m8_genesets<-read_gmt('gene_sets/m8.all.v2024.1.Mm.symbols.gmt.txt')
mh_genesets<-read_gmt('gene_sets/mh.all.v2024.1.Mm.symbols.gmt.txt')
gobp_genesets<-read_gmt('gene_sets/m5.go.bp.v2024.1.Mm.symbols.gmt.txt')
gocc_genesets<-read_gmt('gene_sets/m5.go.cc.v2024.1.Mm.symbols.gmt.txt')
gomf_genesets<-read_gmt('gene_sets/m5.go.mf.v2024.1.Mm.symbols.gmt.txt')
reactome_genesets<-read_gmt('gene_sets/m2.cp.reactome.v2024.1.Mm.symbols.gmt.txt')

Reactome_Regulation_Endogenous_Retroelement<-reactome_genesets["REACTOME_REGULATION_OF_ENDOGENOUS_RETROELEMENTS"]
write.table(Reactome_Regulation_Endogenous_Retroelement, file = "Reactome_Regulation_Endogenous_Retroelement.txt", 
            quote = FALSE, 
            sep = "\t", 
            row.names = FALSE, 
            col.names = FALSE)

```

# b. LT-HSC YA vs YC GSEA

# -> hallmark genesets
```{R}

LT_HSC_YAvsYCWilcox.ordered<-YAvsYCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_YAvsYCWilcox.ordered[order(LT_HSC_YAvsYCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_YAvsYCWilcox.ranks<- LT_HSC_YAvsYCWilcox.ordered$avg_log2FC
names(LT_HSC_YAvsYCWilcox.ranks)<-rownames(LT_HSC_YAvsYCWilcox.ordered)

LT_HSC_YA_vs_YC_mh<- fgseaMultilevel(pathways = mh_genesets, 
                  stats =LT_HSC_YAvsYCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T)

# Extract the number of genes in the leading edge for each gene set
LT_HSC_YA_vs_YC_mh$numLeadingEdgeGenes <- sapply(LT_HSC_YA_vs_YC_mh$leadingEdge, length)

LT_HSC_YA_vs_YC_mh$geneRatio<-LT_HSC_YA_vs_YC_mh$numLeadingEdgeGenes/LT_HSC_YA_vs_YC_mh$size

# View the results with the number of genes in the leading edge
LT_HSC_YA_vs_YC_mh<-LT_HSC_YA_vs_YC_mh[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

  
  xlsx::write.xlsx(LT_HSC_YA_vs_YC_mh, "2025-02-18_Hallmark_LT-HSC_YAvsYC.xlsx")

# Sort by geneRatio
LT_HSC_YA_vs_YC_mh_sorted <- LT_HSC_YA_vs_YC_mh %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_YA_vs_YC_mh_reordered<-ggplot(head(LT_HSC_YA_vs_YC_mh_sorted , 17), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="Hallmarks in LT-HSC YA vs YC") 

  ggsave("2025-02-18_LT-HSC_Hallmarks_YA-vs-YC.pdf",  LT_HSC_YA_vs_YC_mh_reordered, width =8, height = 8, units = c("in"))

```

# -> GO Biological processes
```{r}
LT_HSC_YAvsYCWilcox.ordered<-YAvsYCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_YAvsYCWilcox.ordered[order(LT_HSC_YAvsYCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_YAvsYCWilcox.ranks<- LT_HSC_YAvsYCWilcox.ordered$avg_log2FC
names(LT_HSC_YAvsYCWilcox.ranks)<-rownames(LT_HSC_YAvsYCWilcox.ordered)

LT_HSC_YA_vs_YC_gobp <- fgseaMultilevel(pathways = gobp_genesets, 
                  stats =LT_HSC_YAvsYCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T) 

# Extract the number of genes in the leading edge for each gene set
LT_HSC_YA_vs_YC_gobp$numLeadingEdgeGenes <- sapply(LT_HSC_YA_vs_YC_gobp$leadingEdge, length)

LT_HSC_YA_vs_YC_gobp$geneRatio<-LT_HSC_YA_vs_YC_gobp$numLeadingEdgeGenes/LT_HSC_YA_vs_YC_gobp$size

# View the results with the number of genes in the leading edge
LT_HSC_YA_vs_YC_gobp<-LT_HSC_YA_vs_YC_gobp[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>% 
  as.data.frame()

  xlsx::write.xlsx(LT_HSC_YA_vs_YC_gobp, "2025-02-18_GO-Biological-Processes_LT-HSC_YAvsYC.xlsx")

   # Sort by geneRatio
LT_HSC_YA_vs_YC_gobp_sorted <- LT_HSC_YA_vs_YC_gobp %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_YA_vs_YC_gobp_reordered<-ggplot(head(LT_HSC_YA_vs_YC_gobp_sorted, 17), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="GO BP in LT-HSC YA vs YC") 

  ggsave("2025-02-18_LT-HSC_GOBP_YA-vs-YC.pdf",  LT_HSC_YA_vs_YC_gobp_reordered, width =8, height = 8, units = c("in"))
  

```

# -> Reactome
```{R}

LT_HSC_YAvsYCWilcox.ordered<-YAvsYCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_YAvsYCWilcox.ordered[order(LT_HSC_YAvsYCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_YAvsYCWilcox.ranks<- LT_HSC_YAvsYCWilcox.ordered$avg_log2FC
names(LT_HSC_YAvsYCWilcox.ranks)<-rownames(LT_HSC_YAvsYCWilcox.ordered)



LT_HSC_YAvsYCWilcox_reactome <- fgseaMultilevel(pathways = reactome_genesets, 
                  stats =LT_HSC_YAvsYCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T) 

# Extract the number of genes in the leading edge for each gene set
LT_HSC_YAvsYCWilcox_reactome$numLeadingEdgeGenes <- sapply(LT_HSC_YAvsYCWilcox_reactome$leadingEdge, length)

LT_HSC_YAvsYCWilcox_reactome$geneRatio<-LT_HSC_YAvsYCWilcox_reactome$numLeadingEdgeGenes/LT_HSC_YAvsYCWilcox_reactome$size

# View the results with the number of genes in the leading edge
LT_HSC_YAvsYCWilcox_reactome<-LT_HSC_YAvsYCWilcox_reactome[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

  xlsx::write.xlsx(LT_HSC_YAvsYCWilcox_reactome, "2025-02-18_Reactome_LT-HSC_YAvsYC.xlsx")
  
  # Sort by geneRatio
LT_HSC_YAvsYCWilcox_reactome_sorted <- LT_HSC_YAvsYCWilcox_reactome %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_YAvsYCWilcox_reactome_reordered<-ggplot(head(LT_HSC_YAvsYCWilcox_reactome_sorted, 50), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  #theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="Reactome in LT-HSC YA vs YC") 
  
  ggsave("2025-02-18_LT-HSC_Reactome_YA-vs-YC.pdf", LT_HSC_YAvsYCWilcox_reactome_reordered, width =12, height = 14, units = c("in"))
  
```

# -> Combine all gene sets
```{r}
#since the hallmark gsea results in only 2 enriched pathways, then we will take all of them

LTHSC_YAvsYC_hallmarks_selected<-LT_HSC_YA_vs_YC_mh
LTHSC_YAvsYC_reactome_selected<-LT_HSC_YAvsYCWilcox_reactome
LT_HSC_YA_vs_YC_gobp_selected<-LT_HSC_YA_vs_YC_gobp

LT_HSC_YA_vs_YC_selected_enrichment<- rbind(LTHSC_YAvsYC_hallmarks_selected, LTHSC_YAvsYC_reactome_selected, LT_HSC_YA_vs_YC_gobp_selected) %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_YA_vs_YC_selected_enrichment_plot<-ggplot(LT_HSC_YA_vs_YC_selected_enrichment, aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red", space = "Lab", aesthetics = "colour") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="LT-HSC YA vs YC Top Enrichment") 
  
  ggsave("2025-02-18_LT-HSC_YA-vs-YC_Top_Enrichment.pdf", LT_HSC_YA_vs_YC_selected_enrichment_plot, width =10, height = 5, units = c("in"))

  
LT_HSC_YA_vs_YC_combined_all_enriched_pathways<- rbind(LT_HSC_YA_vs_YC_mh, LT_HSC_YAvsYCWilcox_reactome, LT_HSC_YA_vs_YC_gobp)

  xlsx::write.xlsx(LT_HSC_YA_vs_YC_combined_all_enriched_pathways, "2025-10-29_LT_HSC_YA_vs_YC_combined_all_enriched_pathways.xlsx")

#### Barplot  
  LT_HSC_YA_vs_YC_selected_enrichment_barplot<-ggplot(LT_HSC_YA_vs_YC_selected_enrichment, aes(reorder(pathway, NES), NES, fill = NES)) +
  geom_col() +
  coord_flip() +
  labs(x = "Pathway", y = "Normalized Enrichment Score (NES)",
       title = "GSEA Results") +
  theme_minimal() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0)

 ggsave("2025-02-18_LT-HSC_YA-vs-YC_Top_Enrichment_BARPLOT.pdf.pdf", LT_HSC_YA_vs_YC_selected_enrichment_barplot, width =6, height = 6, units = c("in"))
  
```


# c. LT-HSC OA vs OC GSEA
# -> M8 genesets
```{R}

LT_HSC_OAvsOCWilcox.ordered<-OAvsOCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_OAvsOCWilcox.ordered[order(LT_HSC_OAvsOCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_OAvsOCWilcox.ranks<- LT_HSC_OAvsOCWilcox.ordered$avg_log2FC
names(LT_HSC_OAvsOCWilcox.ranks)<-rownames(LT_HSC_OAvsOCWilcox.ordered)


LT_HSC_OA_vs_OC_M8 <- fgseaMultilevel(pathways = m8_genesets, 
                  stats =LT_HSC_OAvsOCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T)


# Extract the number of genes in the leading edge for each gene set
LT_HSC_OA_vs_OC_M8$numLeadingEdgeGenes <- sapply(LT_HSC_OA_vs_OC_M8$leadingEdge, length)

LT_HSC_OA_vs_OC_M8$geneRatio<-LT_HSC_OA_vs_OC_M8$numLeadingEdgeGenes/LT_HSC_OA_vs_OC_M8$size

# View the results with the number of genes in the leading edge
LT_HSC_OA_vs_OC_M8<-LT_HSC_OA_vs_OC_M8[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>% as.data.frame()

  
  xlsx::write.xlsx(LT_HSC_OA_vs_OC_M8, "2025-02-18_M8_LT-HSC_OAvsOC.xlsx")

 LT_HSC_OAvsOC_M8<-ggplot(head(LT_HSC_OA_vs_OC_M8, 20), aes(y = pathway, x = NES, size = geneRatio)) + 
  geom_point(aes(color = -log10(padj), size= geneRatio)) +
   scale_color_gradient(name = waiver(), low = "blue", high = "red", space = "Lab", aesthetics = "colour")+ labs(x="Normalized Enrichment Score (NES)", y="Pathways",
       title="Tabula Murris in LT-HSC OA vs OC")

  ggsave("2025-01-24_LT-HSC_Tabula-Murris_OA-vs-OC.pdf", LT_HSC_OAvsOC_M8, width =12, height = 8, units = c("in"))


```

# -> hallmark genesets
```{R}

LT_HSC_OAvsOCWilcox.ordered<-OAvsOCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_OAvsOCWilcox.ordered[order(LT_HSC_OAvsOCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_OAvsOCWilcox.ranks<- LT_HSC_OAvsOCWilcox.ordered$avg_log2FC
names(LT_HSC_OAvsOCWilcox.ranks)<-rownames(LT_HSC_OAvsOCWilcox.ordered)

LT_HSC_OA_vs_OC_mh<- fgseaMultilevel(pathways = mh_genesets, 
                  stats =LT_HSC_OAvsOCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T)

# Extract the number of genes in the leading edge for each gene set
LT_HSC_OA_vs_OC_mh$numLeadingEdgeGenes <- sapply(LT_HSC_OA_vs_OC_mh$leadingEdge, length)

LT_HSC_OA_vs_OC_mh$geneRatio<-LT_HSC_OA_vs_OC_mh$numLeadingEdgeGenes/LT_HSC_OA_vs_OC_mh$size

# View the results with the number of genes in the leading edge
LT_HSC_OA_vs_OC_mh<-LT_HSC_OA_vs_OC_mh[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>% 
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

  
  xlsx::write.xlsx(LT_HSC_OA_vs_OC_mh, "2025-02-18_Hallmark_LT-HSC_OAvsOC.xlsx")

# Sort by geneRatio
LT_HSC_OA_vs_OC_mh_sorted <- LT_HSC_OA_vs_OC_mh %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_OA_vs_OC_mh_reordered<-ggplot(head(LT_HSC_OA_vs_OC_mh_sorted , 17), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="Hallmarks in LT-HSC OA vs OC") 

  ggsave("2025-02-18_LT-HSC_Hallmarks_OA-vs-OC.pdf",  LT_HSC_OA_vs_OC_mh_reordered, width =8, height = 8, units = c("in"))

```

# -> GO Biological processes
```{r}
LT_HSC_OAvsOCWilcox.ordered<-OAvsOCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_OAvsOCWilcox.ordered[order(LT_HSC_OAvsOCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_OAvsOCWilcox.ranks<- LT_HSC_OAvsOCWilcox.ordered$avg_log2FC
names(LT_HSC_OAvsOCWilcox.ranks)<-rownames(LT_HSC_OAvsOCWilcox.ordered)

LT_HSC_OA_vs_OC_gobp <- fgseaMultilevel(pathways = gobp_genesets, 
                  stats =LT_HSC_OAvsOCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T) 

# Extract the number of genes in the leading edge for each gene set
LT_HSC_OA_vs_OC_gobp$numLeadingEdgeGenes <- sapply(LT_HSC_OA_vs_OC_gobp$leadingEdge, length)

LT_HSC_OA_vs_OC_gobp$geneRatio<-LT_HSC_OA_vs_OC_gobp$numLeadingEdgeGenes/LT_HSC_OA_vs_OC_gobp$size

# View the results with the number of genes in the leading edge
LT_HSC_OA_vs_OC_gobp<-LT_HSC_OA_vs_OC_gobp[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>% 
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

  xlsx::write.xlsx(LT_HSC_OA_vs_OC_gobp, "2025-02-18_GO-Biological-Processes_LT-HSC_OAvsOC.xlsx")

   # Sort by geneRatio
LT_HSC_OA_vs_OC_gobp_sorted <- LT_HSC_OA_vs_OC_gobp %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_OA_vs_OC_gobp_reordered<-ggplot(head(LT_HSC_OA_vs_OC_gobp_sorted, 17), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="GO BP in LT-HSC OA vs OC") 

  ggsave("2025-02-18_LT-HSC_GOBP_OA-vs-OC.pdf",  LT_HSC_OA_vs_OC_gobp_reordered, width =10, height = 8, units = c("in"))

```

```{r}
LT_HSC_OA_vs_OC_aging_signature<- fgseaMultilevel(pathways = aging_signature_HSC, 
                  stats =LT_HSC_OAvsOCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T) 

```

# -> Reactome
```{R}

LT_HSC_OAvsOCWilcox.ordered<-OAvsOCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

LT_HSC_OAvsOCWilcox.ordered[order(LT_HSC_OAvsOCWilcox.ordered$avg_log2FC, decreasing = T),]
 
LT_HSC_OAvsOCWilcox.ranks<- LT_HSC_OAvsOCWilcox.ordered$avg_log2FC
names(LT_HSC_OAvsOCWilcox.ranks)<-rownames(LT_HSC_OAvsOCWilcox.ordered)


LT_HSC_OA_vs_OC_reactome <- fgseaMultilevel(pathways = reactome_genesets, 
                  stats =LT_HSC_OAvsOCWilcox.ranks,
                  minSize=5,
                  maxSize=1000, 
                  nPermSimple = 10000,
                  gseaParam = T) 

# Extract the number of genes in the leading edge for each gene set
LT_HSC_OA_vs_OC_reactome$numLeadingEdgeGenes <- sapply(LT_HSC_OA_vs_OC_reactome$leadingEdge, length)

LT_HSC_OA_vs_OC_reactome$geneRatio<-LT_HSC_OA_vs_OC_reactome$numLeadingEdgeGenes/LT_HSC_OA_vs_OC_reactome$size

# View the results with the number of genes in the leading edge
LT_HSC_OA_vs_OC_reactome<-LT_HSC_OA_vs_OC_reactome[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size","numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

  xlsx::write.xlsx(LT_HSC_OA_vs_OC_reactome, "2025-02-18_Reactome_LT-HSC_OAvsOC.xlsx")
  
  # Sort by geneRatio
LT_HSC_OA_vs_OC_reactome_sorted <- LT_HSC_OA_vs_OC_reactome %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_OAvsOC_reactome_reordered<-ggplot(head(LT_HSC_OA_vs_OC_reactome_sorted, 50), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  #theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="Reactome in LT-HSC OA vs OC") 
  
  ggsave("2025-02-18_LT-HSC_Reactome_OA-vs-OC.pdf", LT_HSC_OAvsOC_reactome_reordered, width =12, height = 14, units = c("in"))
  
```


# -> Combine some selected gsea results from hallmarks, go bp and reactome and plot them for LT-HSC OA vs OC
```{R}
#since the hallmark gsea results in only 2 enriched pathways, then we will take all of them
LTHSC_OAvsOC_enrichment_Results<-rbind(LT_HSC_OA_vs_OC_mh, LT_HSC_OA_vs_OC_reactome, LT_HSC_OA_vs_OC_gobp)

  xlsx::write.xlsx(LTHSC_OAvsOC_enrichment_Results, "2025-10-29_LT_HSC_OA_vs_OC_combined_all_enriched_pathways.xlsx")


LTHSC_OAvsOC_hallmarks_selected<-LT_HSC_OA_vs_OC_mh
LTHSC_OAvsOC_reactome_selected<-LT_HSC_OA_vs_OC_reactome[1:20,]
LT_HSC_OA_vs_OC_gobp_selected<-LT_HSC_OA_vs_OC_gobp

LT_HSC_OA_vs_OC_selected_enrichment<- rbind(LTHSC_OAvsOC_hallmarks_selected, LTHSC_OAvsOC_reactome_selected, LT_HSC_OA_vs_OC_gobp_selected) %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
LT_HSC_OA_vs_OC_selected_enrichment_plot<-ggplot(LT_HSC_OA_vs_OC_selected_enrichment, aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red", space = "Lab", aesthetics = "colour") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="LT-HSC OA vs OC Top Enrichment") 
  
  ggsave("2025-02-18_LT-HSC_OA-vs-OC_Top_Enrichment.pdf", LT_HSC_OA_vs_OC_selected_enrichment_plot, width =12, height = 12, units = c("in"))

### Create barplot of GSEA OA vs OC LT-HSC
  
LT_HSC_OA_vs_OC_selected_enrichment_barplot<-ggplot(LT_HSC_OA_vs_OC_selected_enrichment, aes(reorder(pathway, NES), NES, fill = NES)) +
  geom_col() +
  coord_flip() +
  labs(x = "Pathway", y = "Normalized Enrichment Score (NES)",
       title = "GSEA Results") +
  theme_minimal() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0)

 ggsave("2025-05-20_LT-HSC_OA-vs-OC_Top_Enrichment_barplot.pdf", LT_HSC_OA_vs_OC_selected_enrichment_barplot, width =12, height = 10, units = c("in"))
```


# -> Selected GSEA to map
```{r}
#### Selected top GSEA pathways
selectedPathway<-c("GOBP_PROTON_MOTIVE_FORCE_DRIVEN_ATP_SYNTHESIS",
          "REACTOME_EPIGENETIC_REGULATION_OF_GENE_EXPRESSION",
          "REACTOME_REGULATION_OF_ENDOGENOUS_RETROELEMENTS",
          "REACTOME_PRC2_METHYLATES_HISTONES_AND_DNA",
          "REACTOME_HDACS_DEACETYLATE_HISTONES",
          "REACTOME_CELLULAR_SENESCENCE",
          "REACTOME_BASE_EXCISION_REPAIR_AP_SITE_FORMATION",
          "REACTOME_DNA_DAMAGE_TELOMERE_STRESS_INDUCED_SENESCENCE",
          "HALLMARK_E2F_TARGETS",
          "REACTOME_CELL_CYCLE",
          "REACTOME_MITOTIC_PROPHASE",
          "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
          "GOBP_NADH_DEHYDROGENASE_COMPLEX_ASSEMBLY")

LT_HSC_OA_vs_OC_selectedPath<-LT_HSC_OA_vs_OC_selected_enrichment %>%
  dplyr::filter(LT_HSC_OA_vs_OC_selected_enrichment$pathway %in% selectedPathway) %>%
  arrange(desc(geneRatio))

# Create dot plot using ggplot2
selectedPathway_LTHSC_OAvsOC<-ggplot(LT_HSC_OA_vs_OC_selectedPath, aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red", space = "Lab", aesthetics = "colour") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="LT-HSC OA vs OC selected Top Enrichment") 
  
  ggsave("2025-02-18_LT-HSC_OA-vs-OC_Top_selected_Enrichment.pdf", selectedPathway_LTHSC_OAvsOC, width =8, height = 8, units = c("in"))
  
```


# -> Combine all gene-set results
```{r}
LT_HSC_OA_vs_OC_combined_all_enriched_pathways<- rbind(LT_HSC_OA_vs_OC_mh, LT_HSC_OA_vs_OC_gobp, LT_HSC_OA_vs_OC_reactome) %>%
  arrange(desc(geneRatio))

  xlsx::write.xlsx(LT_HSC_OA_vs_OC_combined_all_enriched_pathways, "2025-02-18_LT_HSC_OA_vs_OC_combined_all_enriched_pathways.xlsx")
 
#### Barplots of aging and immune pathways
inflammaging<-c("REACTOME_DNA_DAMAGE_TELOMERE_STRESS_INDUCED_SENESCENCE",
                 "REACTOME_INHIBITION_OF_DNA_RECOMBINATION_AT_TELOMERE",
                 "REACTOME_CELLULAR_SENESCENCE",
                 "REACTOME_PROCESSING_OF_DNA_DOUBLE_STRAND_BREAK_ENDS",
                 "REACTOME_DNA_REPAIR", 
                 "REACTOME_G2_M_DNA_DAMAGE_CHECKPOINT", 
                "HALLMARK_TNFA_SIGNALING_VIA_NFKB",
                  "REACTOME_ANTIMICROBIAL_PEPTIDES", 
                  "GOBP_RESPONSE_TO_BACTERIUM",
                  "GOBP_LEUKOCYTE_CHEMOTAXIS",
                  "REACTOME_INTERLEUKIN_7_SIGNALING",
                  "GOBP_HUMORAL_IMMUNE_RESPONSE",
                  "GOBP_DEFENSE_RESPONSE_TO_BACTERIUM",
                  "GOBP_MYELOID_LEUKOCYTE_MIGRATION",
                  "REACTOME_REGULATION_OF_TLR_BY_ENDOGENOUS_LIGAND",
                  "REACTOME_INNATE_IMMUNE_SYSTEM",
                  "GOBP_RESPONSE_TO_BACTERIUM",
                  "GOBP_LEUKOCYTE_MIGRATION_INVOLVED_IN_INFLAMMATORY_RESPONSE",
                  "REACTOME_ROS_AND_RNS_PRODUCTION_IN_PHAGOCYTES",
                  "REACTOME_FCGR_ACTIVATION",
                  "GOBP_NEUTROPHIL_CHEMOTAXIS",
                  "REACTOME_NEUTROPHIL_DEGRANULATION"
                 )

LT_HSC_OA_vs_OC_inflammaging_pathway<-LT_HSC_OA_vs_OC_combined_all_enriched_pathways %>%
  dplyr::filter(LT_HSC_OA_vs_OC_combined_all_enriched_pathways$pathway %in% inflammaging) %>%
  arrange(desc(geneRatio))
  
inflammaging_LTHSC_OAvsOC<-ggplot(LT_HSC_OA_vs_OC_inflammaging_pathway, aes(reorder(pathway, NES), NES, fill = NES)) +
  geom_col() +
  coord_flip() +
  labs(x = "Pathway", y = "Normalized Enrichment Score (NES)",
       title = "GSEA Results") +
  theme_minimal() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0)

 ggsave("2025-03-01_LT-HSC_OA-vs-OC_InflammagingPathways.pdf", inflammaging_LTHSC_OAvsOC, width =10, height = 8, units = c("in"))

 
LT_HSC_OA_vs_OC_inflammaging_dotplot<-ggplot(LT_HSC_OA_vs_OC_inflammaging_pathway, aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red", space = "Lab", aesthetics = "colour") +
  theme_minimal() +
  labs(x = "Gene Ratio", y = "Pathway", color = "NES", size = "-log10(padj)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="LT-HSC OA vs OC Top Enrichment") 
  
  ggsave("2025-02-18_LT-HSC_OA-vs-OC_Inflammaging-dotplot.pdf", LT_HSC_OA_vs_OC_inflammaging_dotplot, width =12, height = 14, units = c("in"))
```

#d. LT-HSC OC vs YC
#-> hallmark gene sets
```{R}
# Filter DEGs: OC vs YC
LT_HSC_OCvsYCWilcox.ordered <- OCvsYCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

# Order by log2FC
LT_HSC_OCvsYCWilcox.ordered <- LT_HSC_OCvsYCWilcox.ordered[order(LT_HSC_OCvsYCWilcox.ordered$avg_log2FC, decreasing = TRUE), ]

# Create ranked gene list
LT_HSC_OCvsYCWilcox.ranks <- LT_HSC_OCvsYCWilcox.ordered$avg_log2FC
names(LT_HSC_OCvsYCWilcox.ranks) <- rownames(LT_HSC_OCvsYCWilcox.ordered)

# Run GSEA with fgseaMultilevel
LT_HSC_OC_vs_YC_mh <- fgseaMultilevel(
  pathways = mh_genesets, 
  stats = LT_HSC_OCvsYCWilcox.ranks,
  minSize = 5,
  maxSize = 1000, 
  nPermSimple = 10000,
  gseaParam = TRUE
)

# Calculate leading edge gene stats
LT_HSC_OC_vs_YC_mh$numLeadingEdgeGenes <- sapply(LT_HSC_OC_vs_YC_mh$leadingEdge, length)
LT_HSC_OC_vs_YC_mh$geneRatio <- LT_HSC_OC_vs_YC_mh$numLeadingEdgeGenes / LT_HSC_OC_vs_YC_mh$size

# Clean up and filter by adjusted p-value
LT_HSC_OC_vs_YC_mh <- LT_HSC_OC_vs_YC_mh[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size", "numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

# Export results to Excel
xlsx::write.xlsx(LT_HSC_OC_vs_YC_mh, "2025-02-18_Hallmark_LT-HSC_OCvsYC.xlsx")

# Sort by gene ratio
LT_HSC_OC_vs_YC_mh_sorted <- LT_HSC_OC_vs_YC_mh %>%
  arrange(desc(geneRatio))

# Dot plot of top 17 enriched gene sets
LT_HSC_OC_vs_YC_mh_reordered <- ggplot(head(LT_HSC_OC_vs_YC_mh_sorted, 17), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(
    title = "Hallmarks in LT-HSC OC vs YC",
    x = "Gene Ratio",
    y = "Pathway",
    color = "NES",
    size = "-log10(padj)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save dot plot
ggsave("2025-02-18_LT-HSC_Hallmarks_OC-vs-YC.pdf", LT_HSC_OC_vs_YC_mh_reordered, width = 8, height = 8, units = "in")

```

#-> GO Biological processes
```{r}
# Filter DEGs for OC vs YC (adjust as needed based on your object)
LT_HSC_OCvsYCWilcox.ordered <- OCvsYCWilcox %>%
  dplyr::filter(p_val <= 0.01)

# Order by log2FC
LT_HSC_OCvsYCWilcox.ordered <- LT_HSC_OCvsYCWilcox.ordered[order(LT_HSC_OCvsYCWilcox.ordered$avg_log2FC, decreasing = TRUE), ]

# Create ranked gene list
LT_HSC_OCvsYCWilcox.ranks <- LT_HSC_OCvsYCWilcox.ordered$avg_log2FC
names(LT_HSC_OCvsYCWilcox.ranks) <- rownames(LT_HSC_OCvsYCWilcox.ordered)

# Run fgseaMultilevel for GO Biological Processes
LT_HSC_OC_vs_YC_gobp <- fgseaMultilevel(
  pathways = gobp_genesets, 
  stats = LT_HSC_OCvsYCWilcox.ranks,
  minSize = 5,
  maxSize = 1000, 
  nPermSimple = 10000,
  gseaParam = TRUE
)

# Count leading edge genes
LT_HSC_OC_vs_YC_gobp$numLeadingEdgeGenes <- sapply(LT_HSC_OC_vs_YC_gobp$leadingEdge, length)

# Compute gene ratios
LT_HSC_OC_vs_YC_gobp$geneRatio <- LT_HSC_OC_vs_YC_gobp$numLeadingEdgeGenes / LT_HSC_OC_vs_YC_gobp$size

# Filter and clean up table
LT_HSC_OC_vs_YC_gobp <- LT_HSC_OC_vs_YC_gobp[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size", "numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame() 

# Export results to Excel
xlsx::write.xlsx(LT_HSC_OC_vs_YC_gobp, "2025-02-18_GO-Biological-Processes_LT-HSC_OCvsYC.xlsx")

# Sort by gene ratio
LT_HSC_OC_vs_YC_gobp_sorted <- LT_HSC_OC_vs_YC_gobp %>%
  arrange(desc(geneRatio))

# Create dot plot of top 17 enriched GO terms
LT_HSC_OC_vs_YC_gobp_reordered <- ggplot(head(LT_HSC_OC_vs_YC_gobp_sorted, 17), aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  theme_minimal() +
  labs(
    x = "Gene Ratio",
    y = "GO Biological Process",
    color = "NES",
    size = "-log10(padj)",
    title = "GO BP in LT-HSC OC vs YC"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save the dot plot
ggsave("2025-02-18_LT-HSC_GOBP_OC-vs-YC.pdf", LT_HSC_OC_vs_YC_gobp_reordered, width = 10, height = 8, units = "in")

```

#-> Reactome
```{r}
# Step 1: Filter DEGs for OC vs YC
LT_HSC_OCvsYCWilcox.ordered <- OCvsYCWilcox %>% 
  dplyr::filter(p_val <= 0.01)

# Step 2: Order by log2FC
LT_HSC_OCvsYCWilcox.ordered <- LT_HSC_OCvsYCWilcox.ordered[order(LT_HSC_OCvsYCWilcox.ordered$avg_log2FC, decreasing = TRUE), ]

# Step 3: Create ranked gene list for GSEA
LT_HSC_OCvsYCWilcox.ranks <- LT_HSC_OCvsYCWilcox.ordered$avg_log2FC
names(LT_HSC_OCvsYCWilcox.ranks) <- rownames(LT_HSC_OCvsYCWilcox.ordered)

# Step 4: Run Reactome GSEA using fgseaMultilevel
LT_HSC_OC_vs_YC_reactome <- fgseaMultilevel(
  pathways = reactome_genesets, 
  stats = LT_HSC_OCvsYCWilcox.ranks,
  minSize = 5,
  maxSize = 1000, 
  nPermSimple = 10000,
  gseaParam = TRUE
)

# Step 5: Add leading edge gene stats
LT_HSC_OC_vs_YC_reactome$numLeadingEdgeGenes <- sapply(LT_HSC_OC_vs_YC_reactome$leadingEdge, length)
LT_HSC_OC_vs_YC_reactome$geneRatio <- LT_HSC_OC_vs_YC_reactome$numLeadingEdgeGenes / LT_HSC_OC_vs_YC_reactome$size

# Step 6: Filter and format final GSEA results
LT_HSC_OC_vs_YC_reactome <- LT_HSC_OC_vs_YC_reactome[, c("pathway", "pval", "padj", "log2err", "ES", "NES", "size", "numLeadingEdgeGenes", "geneRatio")] %>%
  as_tibble() %>%
  arrange(padj) %>%
  dplyr::filter(padj <= 0.05) %>%
  dplyr::filter(abs(NES) >= 1.5) %>%
  as.data.frame()

# Step 7: Export results to Excel
xlsx::write.xlsx(LT_HSC_OC_vs_YC_reactome, "2025-02-18_Reactome_LT-HSC_OCvsYC.xlsx")

# Step 8: Sort by gene ratio
LT_HSC_OC_vs_YC_reactome_sorted <- LT_HSC_OC_vs_YC_reactome %>%
  arrange(desc(geneRatio))

# Step 9: Create dot plot of top 50 pathways
LT_HSC_OCvsYC_reactome_reordered <- ggplot(head(LT_HSC_OC_vs_YC_reactome_sorted, 50), 
                                            aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red") +
  #theme_minimal() +
  labs(
    title = "Reactome in LT-HSC OC vs YC",
    x = "Gene Ratio",
    y = "Reactome Pathway",
    color = "NES",
    size = "-log10(padj)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Step 10: Save the dot plot
ggsave("2025-02-18_LT-HSC_Reactome_OC-vs-YC.pdf", 
       LT_HSC_OCvsYC_reactome_reordered, 
       width = 12, height = 14, units = "in")

```

#-> combine all the results
```{r}
# Rename individual enrichment sets for OC vs YC
LTHSC_OCvsYC_hallmarks_selected <- LT_HSC_OC_vs_YC_mh
LTHSC_OCvsYC_reactome_selected <- LT_HSC_OC_vs_YC_reactome[1:10,]
LT_HSC_OC_vs_YC_gobp_selected <- LT_HSC_OC_vs_YC_gobp[1:10,]

# Combine enrichment sets
LT_HSC_OC_vs_YC_selected_enrichment <- rbind(
  LTHSC_OCvsYC_hallmarks_selected,
  LTHSC_OCvsYC_reactome_selected,
  LT_HSC_OC_vs_YC_gobp_selected
) %>%
  arrange(desc(geneRatio))

# Create dot plot
LT_HSC_OC_vs_YC_selected_enrichment_plot <- ggplot(LT_HSC_OC_vs_YC_selected_enrichment, 
                                                    aes(x = geneRatio, y = reorder(pathway, geneRatio))) +
  geom_point(aes(size = -log10(padj), color = NES), show.legend = TRUE) +
  scale_color_gradient(low = "blue", high = "red", space = "Lab", aesthetics = "colour") +
  theme_minimal() +
  labs(
    title = "LT-HSC OC vs YC Top Enrichment",
    x = "Gene Ratio",
    y = "Pathway",
    color = "NES",
    size = "-log10(padj)"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Save plot
ggsave(
  "2025-02-18_LT-HSC_OC-vs-YC_Top_Enrichment.pdf",
  LT_HSC_OC_vs_YC_selected_enrichment_plot,
  width = 12, height = 12, units = "in"
)

# Combine all enriched pathways for export
LT_HSC_OC_vs_YC_combined_all_enriched_pathways <- rbind(
  LT_HSC_OC_vs_YC_mh,
  LT_HSC_OC_vs_YC_reactome, 
  LT_HSC_OC_vs_YC_gobp
) 
# Export to Excel
xlsx::write.xlsx(
  LT_HSC_OC_vs_YC_combined_all_enriched_pathways,
  "2025-10-29_LT_HSC_OC_vs_YC_combined_all_enriched_pathways.xlsx"
)

```

##e. LT-HSC OA vs YC
## -> hallmark gene sets



#11. Volcano plots of transposable elements
# Volcano Plot Visualizierung von LT-HSC OA gegen OC
```{r}
library(EnhancedVolcano)
# Beräumung von TE Identität nach ihre Classes, Familie, und die Spezies
TE<-OAvsOCWilcox[grep("gene-id", rownames(OAvsOCWilcox)), ] # Nehme die TE-Identität von DEN-Ergebnissen von LT-HSC OA gegen OC

# TE-Spezies Extraktierung von ihre eigene IDs
species<-stringi::stri_extract(TE$genes, regex = ".*(?=;)") %>% 
  stringi::stri_extract(regex = "(?<=gene-id).*") %>%
  as.vector()

# TE-Classes Extraktierung von ihre eigene IDs
Classes<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = ".*(?=/)") %>%
  replace_na("DNA")
# Hier habt ich realizieren das manche DNA TE hat keine / Symbole um das machen die manche TE hat keine Klasse. Hier muss ich die NA nach DNA andern

# TE-Familie Extraktierung von ihre eigene IDs
Families<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = "(?<=/).*")

DET_LTHSC_OAvsOC<-cbind(TE, species, Families, Classes)

species<-EnhancedVolcano(DET_LTHSC_OAvsOC,
    lab = DET_LTHSC_OAvsOC$species,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-2, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE species in OA vs OC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.3,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


families<-EnhancedVolcano(DET_LTHSC_OAvsOC,
    lab = DET_LTHSC_OAvsOC$Families,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-2, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE families in OA vs OC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.3,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

classes<-EnhancedVolcano(DET_LTHSC_OAvsOC,
    lab = DET_LTHSC_OAvsOC$Classes,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-2, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE classes in OA vs OC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.3,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

print(species)
print(families)
print(classes)

ggsave("TE_volcano_species_OAvsOC.pdf", print(species), width = 7, height = 5)
ggsave("TE_volcano_families_OAvsOC.pdf", print(families), width = 7, height = 5)
ggsave("TE_volcano_classes_OAvsOC.pdf", print(classes), width = 7, height = 5)
```


# Volcano Plot Visualizierung von LT-HSC OA gegen YC
```{r}
# Beräumung von TE Identität nach ihre Classes, Familie, und die Spezies
TE<-OAvsYCWilcox[grep("gene-id", rownames(OAvsYCWilcox)), ] # Nehme die TE-Identität von DEN-Ergebnissen von LT-HSC OA gegen YC

# TE-Spezies Extraktierung von ihre eigene IDs
species<-stringi::stri_extract(TE$genes, regex = ".*(?=;)") %>% 
  stringi::stri_extract(regex = "(?<=gene-id).*") %>%
  as.vector()

# TE-Classes Extraktierung von ihre eigene IDs
Classes<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = ".*(?=/)") %>%
  replace_na("DNA")
# Hier habt ich realizieren das manche DNA TE hat keine / Symbole um das machen die manche TE hat keine Klasse. Hier muss ich die NA nach DNA andern

# TE-Familie Extraktierung von ihre eigene IDs
Families<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = "(?<=/).*")

DET_LTHSC_OAvsYC<-cbind(TE, species, Families, Classes)

species<-EnhancedVolcano(DET_LTHSC_OAvsYC,
    lab = DET_LTHSC_OAvsYC$species,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE species in OA vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


families<-EnhancedVolcano(DET_LTHSC_OAvsYC,
    lab = DET_LTHSC_OAvsYC$Families,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE families in OA vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

classes<-EnhancedVolcano(DET_LTHSC_OAvsYC,
    lab = DET_LTHSC_OAvsYC$Classes,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE classes in OA vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

print(species)
print(families)
print(classes)
```


# Volcano Plot Visualizierung von LT-HSC OC gegen YC
```{r}
# Beräumung von TE Identität nach ihre Classes, Familie, und die Spezies
TE<-OCvsYCWilcox[grep("gene-id", rownames(OCvsYCWilcox)), ] # Nehme die TE-Identität von DEN-Ergebnissen von LT-HSC OC gegen YC

# TE-Spezies Extraktierung von ihre eigene IDs
species<-stringi::stri_extract(TE$genes, regex = ".*(?=;)") %>% 
  stringi::stri_extract(regex = "(?<=gene-id).*") %>%
  as.vector()

# TE-Classes Extraktierung von ihre eigene IDs
Classes<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = ".*(?=/)") %>%
  replace_na("DNA")
# Hier habt ich realizieren das manche DNA TE hat keine / Symbole um das machen die manche TE hat keine Klasse. Hier muss ich die NA nach DNA andern

# TE-Familie Extraktierung von ihre eigene IDs
Families<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = "(?<=/).*")

DET_LTHSC_OCvsYC<-cbind(TE, species, Families, Classes)

species<-EnhancedVolcano(DET_LTHSC_OCvsYC,
    lab = DET_LTHSC_OCvsYC$species,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE species in OC vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.7,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


families<-EnhancedVolcano(DET_LTHSC_OCvsYC,
    lab = DET_LTHSC_OCvsYC$Families,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE families in OC vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.7,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

classes<-EnhancedVolcano(DET_LTHSC_OCvsYC,
    lab = DET_LTHSC_OCvsYC$Classes,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE classes in OC vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.7,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

print(species)
print(families)
print(classes)
```


# Volcano Plot Visualizierung von LT-HSC YA gegen YC
```{r}
# Beräumung von TE Identität nach ihre Classes, Familie, und die Spezies
TE<-YAvsYCWilcox[grep("gene-id", rownames(YAvsYCWilcox)), ] # Nehme die TE-Identität von DEN-Ergebnissen von LT-HSC OA gegen YC

# TE-Spezies Extraktierung von ihre eigene IDs
species<-stringi::stri_extract(TE$genes, regex = ".*(?=;)") %>% 
  stringi::stri_extract(regex = "(?<=gene-id).*") %>%
  as.vector()

# TE-Classes Extraktierung von ihre eigene IDs
Classes<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = ".*(?=/)") %>%
  replace_na("DNA")
# Hier habt ich realizieren das manche DNA TE hat keine / Symbole um das machen die manche TE hat keine Klasse. Hier muss ich die NA nach DNA andern

# TE-Familie Extraktierung von ihre eigene IDs
Families<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = "(?<=/).*")

DET_LTHSC_YAvsYC<-cbind(TE, species, Families, Classes)

species<-EnhancedVolcano(DET_LTHSC_YAvsYC,
    lab = DET_LTHSC_YAvsYC$species,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-2, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE species in YA vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.3,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


families<-EnhancedVolcano(DET_LTHSC_YAvsYC,
    lab = DET_LTHSC_YAvsYC$Families,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-2, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE families in YA vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.3,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

classes<-EnhancedVolcano(DET_LTHSC_YAvsYC,
    lab = DET_LTHSC_YAvsYC$Classes,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-2, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE classes in YA vs YC LT-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.3,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

print(species)
print(families)
print(classes)
ggsave("TE_volcano_species_YAvsYC.pdf", print(species), width = 7, height = 5)
ggsave("TE_volcano_families_YAvsYC.pdf", print(families), width = 7, height = 5)
ggsave("TE_volcano_classes_YAvsYC.pdf", print(classes), width = 7, height = 5)
```


# Volcano Plot Visualizierung von MPP3 YA gegen YC
```{r}
# Beräumung von TE Identität nach ihre Classes, Familie, und die Spezies
TE<-MPP3_YAvsYCWilcox[grep("gene-id", rownames(MPP3_YAvsYCWilcox)), ] # Nehme die TE-Identität von DEN-Ergebnissen von MPP3 YA gegen YC

# TE-Spezies Extraktierung von ihre eigene IDs
species<-stringi::stri_extract(TE$genes, regex = ".*(?=;)") %>% 
  stringi::stri_extract(regex = "(?<=gene-id).*") %>%
  as.vector()

# TE-Classes Extraktierung von ihre eigene IDs
Classes<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = ".*(?=/)") %>%
  replace_na("DNA")
# Hier habt ich realizieren das manche DNA TE hat keine / Symbole um das machen die manche TE hat keine Klasse. Hier muss ich die NA nach DNA andern

# TE-Familie Extraktierung von ihre eigene IDs
Families<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = "(?<=/).*")

DET_MPP3_YAvsYC<-cbind(TE, species, Families, Classes)

species<-EnhancedVolcano(DET_MPP3_YAvsYC,
    lab = DET_MPP3_YAvsYC$species,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE species in YA vs YC MPP3',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


families<-EnhancedVolcano(DET_MPP3_YAvsYC,
    lab = DET_MPP3_YAvsYC$Families,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE families in YA vs YC MPP3-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

classes<-EnhancedVolcano(DET_MPP3_YAvsYC,
    lab = DET_MPP3_YAvsYC$Classes,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE classes in YA vs YC MPP3',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

print(species)
print(families)
print(classes)
```



# Volcano Plot Visualizierung von MPP3 OA gegen OC
```{r}
# Beräumung von TE Identität nach ihre Klasse, Familie, und die Spezies
TE<-MPP3_OAvsOCWilcox[grep("gene-id", rownames(MPP3_OAvsOCWilcox)), ] # Nehme die TE-Identität von DE Analysisergebnisse von MPP3 gruppe YA gegen YC 

# TE-Spezies Extraktion von jede TE gene IDs
species<-stringi::stri_extract(TE$genes, regex = ".*(?=;)") %>% 
  stringi::stri_extract(regex = "(?<=gene-id).*") %>%
  as.vector()

# TE-Classes Extraktion von jede TE gene IDs
Classes<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = ".*(?=/)") %>%
  replace_na("DNA")
# Hier habt ich realizieren das manche DNA TE hat keine / Symbole um das machen die manche TE hat keine Klasse. Hier muss ich die "NA" nach DNA andern

# TE-Familie Extraktierung von ihre eigene IDs
Families<-stringi::stri_extract(TE$genes, regex = "(?<=transcript-id).*") %>%
  stringi::stri_extract(regex = "(?<=/).*")

DET_MPP3_OAvsOC<-cbind(TE, species, Families, Classes)

species<-EnhancedVolcano(DET_MPP3_OAvsOC,
    lab = DET_MPP3_OAvsOC$species,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE species in OA vs OC MPP3',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)


families<-EnhancedVolcano(DET_MPP3_OAvsOC,
    lab = DET_MPP3_YAvsYC$Families,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE families in OA vs OC MPP3-HSC',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

classes<-EnhancedVolcano(DET_MPP3_OAvsOC,
    lab = DET_MPP3_YAvsYC$Classes,
    x = 'avg_log2FC',
    y = 'p_val_adj',
     xlim =c(-1, 2),
    ylim = c(0, 80),
    title = 'Differential expressed TE classes in OA vs OC MPP3',
    xlab = bquote(~Log[2]~ 'fold change'),
    pCutoff = 0.05,
    FCcutoff = 0.1,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 1.0,
    labSize = 2,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log2FC','p-value',
      'p-value & Log2FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    drawConnectors = TRUE,
    #widthConnectors = 0.5,
    max.overlaps = 100,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)

print(species)
print(families)
print(classes)
```

# Create barplot of Differential TE
```{r}
# extract Differentially expressed TE based on padj 0.05
sigTEs_LTHSC_OAvsOC <- DET_LTHSC_OAvsOC %>%
  dplyr::filter(p_val_adj <= 0.05, abs(avg_log2FC) >= 0.3)
sigTEs_LTHSC_OAvsOC$expression <- "NA"
sigTEs_LTHSC_OAvsOC$comparison <- "OAvsOC"

upTE_oa_vs_oc<-sigTEs_LTHSC_OAvsOC %>%
  dplyr::filter(avg_log2FC >= 0.3)
upTE_oa_vs_oc<-upTE_oa_vs_oc$species

downTE_oa_vs_oc<- sigTEs_LTHSC_OAvsOC %>%
  dplyr::filter(avg_log2FC <= -0.3)
downTE_oa_vs_oc<-downTE_oa_vs_oc$species

# Assign 'expression' based on membership in the TE lists
sigTEs_LTHSC_OAvsOC$expression <- ifelse(sigTEs_LTHSC_OAvsOC$species %in% downTE_oa_vs_oc, "downregulated",
                        ifelse(sigTEs_LTHSC_OAvsOC$species %in% upTE_oa_vs_oc, "upregulated", NA))  # or "unknown"

sigTEs_LTHSC_OAvsOC <- sigTEs_LTHSC_OAvsOC %>%
  mutate(across(c(species, Families, Classes, expression), 
                ~ trimws(gsub('"', '', .))))


sigTEs_LTHSC_OAvsOC <- sigTEs_LTHSC_OAvsOC[, c("species", "Families", "Classes", "expression", "p_val", "avg_log2FC", "pct.1", "pct.2", "p_val_adj", "comparison")]

##### YA vs YC LT-HSC
sigTEs_LTHSC_YAvsYC <- DET_LTHSC_YAvsYC %>%
  dplyr::filter(p_val_adj <= 0.05, abs(avg_log2FC) >= 0.3)
sigTEs_LTHSC_YAvsYC$expression<- "NA"
sigTEs_LTHSC_YAvsYC$comparison<- "YAvsYC"

upTE_ya_vs_yc<-sigTEs_LTHSC_YAvsYC %>%
  dplyr::filter(avg_log2FC >= 0.3)
upTE_ya_vs_yc<-upTE_ya_vs_yc$species

downTE_ya_vs_yc <-sigTEs_LTHSC_YAvsYC %>%
  dplyr::filter(avg_log2FC <= -0.3)
downTE_ya_vs_yc <- downTE_ya_vs_yc$species

# Assign 'expression' based on membership in the TE lists
sigTEs_LTHSC_YAvsYC$expression <- ifelse(sigTEs_LTHSC_YAvsYC$species %in% downTE_ya_vs_yc, "downregulated", 
                                         ifelse(sigTEs_LTHSC_YAvsYC$species %in% upTE_ya_vs_yc, "upregulated", NA))

# Assign 'expression' based on membership in the TE lists
sigTEs_LTHSC_YAvsYC$expression <- ifelse(sigTEs_LTHSC_YAvsYC$species %in% downTE_ya_vs_yc, "downregulated",
                                  ifelse(sigTEs_LTHSC_YAvsYC$species %in% upTE_ya_vs_yc, "upregulated", NA))  # or "unknown"

# Remove quotes from selected columns
sigTEs_LTHSC_YAvsYC <- sigTEs_LTHSC_YAvsYC %>%
  mutate(across(c(species, Families, Classes, expression), 
                ~ trimws(gsub('"', '', .))))


# Subset to selected columns
sigTEs_LTHSC_YAvsYC <- sigTEs_LTHSC_YAvsYC[, c("species", "Families", "Classes", "expression", 
                                               "p_val", "avg_log2FC", "pct.1", "pct.2", "p_val_adj", "comparison")]

sigTEs<-rbind(sigTEs_LTHSC_YAvsYC, sigTEs_LTHSC_OAvsOC)

```

```{r}
library(ggplot2)
library(dplyr)
sigTEs$comparison <- factor(sigTEs$comparison, levels = c("YAvsYC", "OAvsOC"))

# Summarize count of TEs per Class, expression type, and comparison
plot_data <- sigTEs %>%
  group_by(Classes, expression, comparison) %>%
  summarise(count = n(), .groups = "drop")

# Define custom colors for expression levels
custom_colors <- c(
  "upregulated" = "#d62728" ,     # red
  "downregulated" = "#1f77b4",   # blue
  "non_change" = "#999999"       # gray (optional if included)
)


# Barplot
te_distribution<-ggplot(plot_data, aes(x = Classes, y = count, fill = expression)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ comparison) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(
    title = "Distribution of TE Classes by Expression and Comparison",
    x = "TE Class",
    y = "Count",
    fill = "Expression"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("DET_distribution.pdf", print(te_distribution), width = 5, height = 5)
```
```{r}
library(ggplot2)
library(dplyr)

# Step 1: Calculate counts and proportions per Class and Comparison
plot_data <- sigTEs %>%
  group_by(comparison, Classes, expression) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(comparison, Classes) %>%
  mutate(proportion = count / sum(count))  # this creates stacked bar proportions

# Step 2: Plot stacked barplot with proportions
ggplot(plot_data, aes(x = comparison, y = proportion, fill = expression)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Classes) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() +
  labs(
    title = "Proportion of Up/Downregulated TEs by Class and Comparison",
    x = "Comparison",
    y = "Proportion",
    fill = "Expression"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# session info
```{r}
sessionInfo()
```

